---
title: "Reproducible Research - Project 1"
autor: Angel Villamizar
date: "24/10/2022"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth: 1
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
library(knitr)
library(dplyr)
library(data.table)
library(ggplot2)

directory=getwd()

workspace=getwd()
workspace=sprintf("%s/.RData",workspace)
```

1.- Code for reading in the dataset and/or processing the data.
```{r echo=TRUE}
fichero <- list.files(directory, pattern = ".zip" )

data <- unzip(fichero, 
      files = NULL, list = FALSE, overwrite = TRUE, junkpaths = FALSE, 
      exdir = directory, unzip = "internal", setTimes = FALSE)

activity <- read.csv(data, sep = ",")

head(activity)
```

2.- Histogram of the total number of steps taken each day.
```{r echo=TRUE}
steps_by_day <- activity %>%
    group_by(date) %>%
    summarize(total = sum(steps))  %>%  as.data.table()

hist(steps_by_day$total, main="Histogram of total number of steps per day", 
     xlab="Total number of steps in a day")

```

3.- Mean and median number of steps taken each day.
```{r echo=TRUE}
summary(steps_by_day)
```

Mean of total number of steps per day is 10766, median is 10765.

4.- Time series plot of the average number of steps taken.
```{r echo=TRUE}
steps_by_interval <- aggregate(steps ~ interval, activity, mean)

plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")

```

5.- The 5-minute interval that, on average, contains the maximum number of steps.
```{r echo=TRUE}
max_steps_row <- which.max(steps_by_interval$steps)

steps_by_interval[max_steps_row, ]
```

The interval 835 has the maximum average value of steps = 206.1698.

6.- Code to describe and show a strategy for imputing missing data.
```{r echo=TRUE}
sum(is.na(activity))
```

Total number of rows with NA’s is 2304.

Replacing NA’s with the mean for that 5-minute interval.
```{r echo=TRUE}
aux_steps <- data.table()
for (i in 1:nrow(activity)) {
  # i <- 2
  activity2 <- activity[i,]
  
  if (is.na(activity2$steps)) {
    
    interv <- activity2[,c("interval")]
  
    mean_interv <- steps_by_interval[steps_by_interval$interval == interv,c("steps")]
    
    activity2$steps <- mean_interv
    
  }else{ activity2$steps <- activity2$steps }
  
  aux_steps <- rbind(aux_steps,activity2)
}

aux_steps2 <- aux_steps
aux_steps2$steps <- round(aux_steps2$steps, 0)
```

7.- Mean and median number of steps taken each day after missing values are imputed.
```{r echo=TRUE}
aux_steps2_steps_by_day <- aggregate(steps ~ date, aux_steps2, sum)

mean(aux_steps2_steps_by_day$steps)

median(aux_steps2_steps_by_day$steps)
```

8.- Histogram of the total number of steps taken each day after missing values are imputed.
```{r echo=TRUE}
df_imputed <- aggregate(steps ~ date, aux_steps2, sum)
head(df_imputed)

hist(df_imputed$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
```

9.- Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r echo=TRUE}
setDF(aux_steps2)
aux_steps2['day'] <- weekdays(as.Date(aux_steps2$date))

aux_steps2$day[aux_steps2$day  %in% c('sábado','domingo') ] <- "weekend"
aux_steps2$day[aux_steps2$day != "weekend"] <- "weekday"

aux_steps2$day <- as.factor(aux_steps2$day)

aux_steps2_interval <- aggregate(steps ~ interval + day, aux_steps2, mean)

qplot(interval, 
      steps, 
      data = aux_steps2_interval, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
  facet_wrap(~ day, ncol = 1)

```

```{r include=FALSE}
save.image(workspace)
```


